---
title: "Universal Cervical Length Screening in Twins: Cost-effectiveness Analysis (Base Case 2/2)"
author: "Yannay Khaikin and Rachel Gladstone"
date: "2024-10-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
library(ggplot2)
```

### Setup

```{r}
# Parameters from "CLS Model_2states_V2.Rmd" required
n_age_max
v_GA_cat
```

### Annual mortality rate

1. Baseline mortality rate 
  + *SOURCE: Human Mortality Database (Canada 2020)*
2. Age-dependent hazard ratios for mortality by gestational age at delivery
  + *SOURCE: Crump Lancet Child Adol Health 2019 (PMID: 30956154), Table 2*

```{r, warning = FALSE}
## Baseline mortality rate (Canada 2020 data) 
mort_data <- read.table("Data/Canada_mort_data.txt", header = TRUE)

# Mortality data for latest available year (2020) and for ages 0-99
v_r_mort <- mort_data %>%
  mutate_if(is.character, as.numeric) %>%
  filter(Year == 2020, Age <= (n_age_max - 1)) %>%
  select(qx) %>%
  as.matrix()

## Hazard ratios of mortality by age group and GA group (Crump 2019)

# Age 0 to <1
hr_AD_28_01 <- 236.32
hr_AD_32_01 <- 32.1
hr_AD_34_01 <- 32.1
hr_AD_37_01 <- 6.75
hr_AD_term_01 <- 1

# Age 1 to 9
length_19 <- 9-1+1
hr_AD_28_19 <- 4.52
hr_AD_32_19 <- 3.26
hr_AD_34_19 <- 3.26
hr_AD_37_19 <- 1.9
hr_AD_term_19 <- 1

# Age 10 to 19 
length_1019 <- 19-10+1
hr_AD_28_1019 <- 1.89
hr_AD_32_1019 <- 1.91
hr_AD_34_1019 <- 1.91
hr_AD_37_1019 <- 1.37 
hr_AD_term_1019 <- 1

# Age 20-29 
length_2029 <- 29-20+1
hr_AD_28_2029 <- 2.12
hr_AD_32_2029 <- 1.42
hr_AD_34_2029 <- 1.42
hr_AD_37_2029 <- 1.35
hr_AD_term_2029 <- 1

# Age 30-45
length_3045 <- 45-30+1
hr_AD_28_3045 <- 2.04
hr_AD_32_3045 <- 1.48
hr_AD_34_3045 <- 1.48
hr_AD_37_3045 <- 1.22
hr_AD_term_3045 <- 1

# Plateau from 46-100 (all resolve to 1)
length_100 <- 100-46
hr_final <- 1

## Mortality age-dependent hazard ratios by GA category # Labels ex. "Very preterm" from the source paper

# Extremely preterm (< 28 wks)
v_hr_AD_28 <- c(hr_AD_28_01, 
                rep(hr_AD_28_19, length_19), 
                rep(hr_AD_28_1019, length_1019),
                rep(hr_AD_28_2029, length_2029), 
                rep(hr_AD_28_3045, length_3045),
                seq(from = hr_AD_28_3045, to = hr_final, length.out = length_100))

# Very preterm (28-33 wks)
v_hr_AD_32 <- c(hr_AD_32_01, 
                rep(hr_AD_32_19, length_19), 
                rep(hr_AD_32_1019, length_1019),
                rep(hr_AD_32_2029, length_2029), 
                rep(hr_AD_32_3045, length_3045),
                seq(from = hr_AD_32_3045, to = hr_final, length.out = length_100))

# Very preterm (28-33 wks)
v_hr_AD_34 <- c(hr_AD_34_01, 
                rep(hr_AD_34_19, length_19), 
                rep(hr_AD_34_1019, length_1019),
                rep(hr_AD_34_2029, length_2029), 
                rep(hr_AD_34_3045, length_3045),
                seq(from = hr_AD_34_3045, to = hr_final, length.out = length_100))

# Late pre-term (34-36)
v_hr_AD_37 <- c(hr_AD_37_01, 
                rep(hr_AD_37_19, length_19), 
                rep(hr_AD_37_1019, length_1019),
                rep(hr_AD_37_2029, length_2029), 
                rep(hr_AD_37_3045, length_3045),
                seq(from = hr_AD_37_3045, to = hr_final, length.out = length_100))

# Term
v_hr_AD_term <- c(hr_AD_term_01, 
                  rep(hr_AD_term_19, length_19), 
                  rep(hr_AD_term_1019, length_1019),
                  rep(hr_AD_term_2029, length_2029), 
                  rep(hr_AD_term_3045, length_3045),
                  seq(from = hr_AD_term_3045, to = hr_final, length.out = length_100))

df_hr <- data.frame(age = c(0:99),
                    hr_28 = v_hr_AD_28,
                    hr_32 = v_hr_AD_32,
                    hr_34 = v_hr_AD_34,
                    hr_37 = v_hr_AD_37,
                    hr_term = v_hr_AD_term)
                    
# Visual check for HR vectors (note the ylim)
ggplot(data = df_hr, aes(x = age)) +
  geom_line(aes(y = hr_28), color = "darkgreen") +
  geom_line(aes(y = hr_32), color = "darkblue") +
  geom_line(aes(y = hr_34), color = "darkred") +
  geom_line(aes(y = hr_37), color = "orange") +
  geom_line(aes(y = hr_term), color = "pink") +
  labs(x = "Age", 
       y = "Hazard ratio of death",
       title = "Age-dependent hazard ratios by GA category") +
  theme_bw()

## Final mortality rate calculation and conversion to probability occurs in the main document "CLS_model_basecase_1.Rmd"
```

### Costs

Annual costs by gestational age at birth.
*SOURCE: Johnston BMC Pediatrics 2014 (PMID: 24708755)*

```{r}
## Vector of annual costs post-discharge
# Df contains inpatient and outpatient costs, as well as standard errors for each
df_c <- read.csv("Data/johnston2014_costs.csv", header = TRUE, skip = 1)

# MODEL INPUTS: Inpatient cost and SE by Johnston GA category and age (Johnston 2014 categories)

# < 28
in_28_0 <- c(mean = df_c[1, 4], se = df_c[1, 5])
in_28_1 <- c(mean = df_c[2, 4], se = df_c[2, 5])
in_28_2 <- c(mean = df_c[3, 4], se = df_c[3, 5])
in_28_3 <- c(mean = df_c[4, 4], se = df_c[4, 5])
in_28_4 <- c(mean = df_c[5, 4], se = df_c[5, 5])
in_28_5 <- c(mean = df_c[6, 4], se = df_c[6, 5])
in_28_6 <- c(mean = df_c[7, 4], se = df_c[7, 5])
in_28_7 <- c(mean = df_c[8, 4], se = df_c[8, 5])
in_28_8 <- c(mean = df_c[9, 4], se = df_c[9, 5])
in_28_9 <- c(mean = df_c[10, 4], se = df_c[10, 5])

# 28-32
in_32_0 <- c(mean = df_c[1, 12], se = df_c[1, 13])
in_32_1 <- c(mean = df_c[2, 12], se = df_c[2, 13])
in_32_2 <- c(mean = df_c[3, 12], se = df_c[3, 13])
in_32_3 <- c(mean = df_c[4, 12], se = df_c[4, 13])
in_32_4 <- c(mean = df_c[5, 12], se = df_c[5, 13])
in_32_5 <- c(mean = df_c[6, 12], se = df_c[6, 13])
in_32_6 <- c(mean = df_c[7, 12], se = df_c[7, 13])
in_32_7 <- c(mean = df_c[8, 12], se = df_c[8, 13])
in_32_8 <- c(mean = df_c[9, 12], se = df_c[9, 13])
in_32_9 <- c(mean = df_c[10, 12], se = df_c[10, 13])

# 33-36
in_36_0 <- c(mean = df_c[1, 20], se = df_c[1, 21])
in_36_1 <- c(mean = df_c[2, 20], se = df_c[2, 21])
in_36_2 <- c(mean = df_c[3, 20], se = df_c[3, 21])
in_36_3 <- c(mean = df_c[4, 20], se = df_c[4, 21])
in_36_4 <- c(mean = df_c[5, 20], se = df_c[5, 21])
in_36_5 <- c(mean = df_c[6, 20], se = df_c[6, 21])
in_36_6 <- c(mean = df_c[7, 20], se = df_c[7, 21])
in_36_7 <- c(mean = df_c[8, 20], se = df_c[8, 21])
in_36_8 <- c(mean = df_c[9, 20], se = df_c[9, 21])
in_36_9 <- c(mean = df_c[10, 20], se = df_c[10, 21])

# Outpatient costs by Johnston GA category and age
# < 28
out_28_0 <- c(mean = df_c[1, 8], se = df_c[1, 9])
out_28_1 <- c(mean = df_c[2, 8], se = df_c[2, 9])
out_28_2 <- c(mean = df_c[3, 8], se = df_c[3, 9])
out_28_3 <- c(mean = df_c[4, 8], se = df_c[4, 9])
out_28_4 <- c(mean = df_c[5, 8], se = df_c[5, 9])
out_28_5 <- c(mean = df_c[6, 8], se = df_c[6, 9])
out_28_6 <- c(mean = df_c[7, 8], se = df_c[7, 9])
out_28_7 <- c(mean = df_c[8, 8], se = df_c[8, 9])
out_28_8 <- c(mean = df_c[9, 8], se = df_c[9, 9])
out_28_9 <- c(mean = df_c[10, 8], se = df_c[10, 9])

# 28-32
out_32_0 <- c(mean = df_c[1, 16], se = df_c[1, 17])
out_32_1 <- c(mean = df_c[2, 16], se = df_c[2, 17])
out_32_2 <- c(mean = df_c[3, 16], se = df_c[3, 17])
out_32_3 <- c(mean = df_c[4, 16], se = df_c[4, 17])
out_32_4 <- c(mean = df_c[5, 16], se = df_c[5, 17])
out_32_5 <- c(mean = df_c[6, 16], se = df_c[6, 17])
out_32_6 <- c(mean = df_c[7, 16], se = df_c[7, 17])
out_32_7 <- c(mean = df_c[8, 16], se = df_c[8, 17])
out_32_8 <- c(mean = df_c[9, 16], se = df_c[9, 17])
out_32_9 <- c(mean = df_c[10, 16], se = df_c[10, 17])

# 33-36
out_36_0 <- c(mean = df_c[1, 24], se = df_c[1, 25])
out_36_1 <- c(mean = df_c[2, 24], se = df_c[2, 25])
out_36_2 <- c(mean = df_c[3, 24], se = df_c[3, 25])
out_36_3 <- c(mean = df_c[4, 24], se = df_c[4, 25])
out_36_4 <- c(mean = df_c[5, 24], se = df_c[5, 25])
out_36_5 <- c(mean = df_c[6, 24], se = df_c[6, 25])
out_36_6 <- c(mean = df_c[7, 24], se = df_c[7, 25])
out_36_7 <- c(mean = df_c[8, 24], se = df_c[8, 25])
out_36_8 <- c(mean = df_c[9, 24], se = df_c[9, 25])
out_36_9 <- c(mean = df_c[10, 24], se = df_c[10, 25])

# Average value at age 10 - to be used for 10-100 for all categories (calculated within the model)
in_10 <- c(mean = mean(in_28_9[1], in_32_9[1], in_36_9[1]))
out_10 <- c(mean = mean(out_28_9[1], out_32_9[1], out_36_9[1]))

# Length from age 10 to 100 inclusive (101 total costs from 0:100)
length_101 <- 101-10 

## Using the individual costs by year, combine to form vectors for Markov model
# Vector of inpatient costs by GA
v_c_in_28 <- c(in_28_0[1], in_28_1[1], in_28_2[1], in_28_3[1], in_28_4[1], in_28_5[1], in_28_6[1], in_28_7[1], in_28_8[1], in_28_9[1], 
               rep(in_10[1], length_101))

v_c_in_32 <- c(in_32_0[1], in_32_1[1], in_32_2[1], in_32_3[1], in_32_4[1], in_32_5[1], in_32_6[1], in_32_7[1], in_32_8[1], in_32_9[1], 
               rep(in_10[1], length_101))

v_c_in_34 <- c(in_36_0[1], in_36_1[1], in_36_2[1], in_36_3[1], in_36_4[1], in_36_5[1], in_36_6[1], in_36_7[1], in_36_8[1], in_36_9[1], 
               rep(in_10[1], length_101))

v_c_in_37 <- c(in_36_0[1], in_36_1[1], in_36_2[1], in_36_3[1], in_36_4[1], in_36_5[1], in_36_6[1], in_36_7[1], in_36_8[1], in_36_9[1], 
               rep(in_10[1], length_101))

v_c_in_term <- c(in_36_0[1], in_36_1[1], in_36_2[1], in_36_3[1], in_36_4[1], in_36_5[1], in_36_6[1], in_36_7[1], in_36_8[1], in_36_9[1], 
                 rep(in_10[1], length_101))

# Vector of outpatient costs by GA
v_c_out_28 <- c(out_28_0[1], out_28_1[1], out_28_2[1], out_28_3[1], out_28_4[1], out_28_5[1], out_28_6[1], out_28_7[1], out_28_8[1], out_28_9[1], 
               rep(out_10[1], length_101))

v_c_out_32 <- c(out_32_0[1], out_32_1[1], out_32_2[1], out_32_3[1], out_32_4[1], out_32_5[1], out_32_6[1], out_32_7[1], out_32_8[1], out_32_9[1], 
               rep(out_10[1], length_101))

v_c_out_34 <- c(out_36_0[1], out_36_1[1], out_36_2[1], out_36_3[1], out_36_4[1], out_36_5[1], out_36_6[1], out_36_7[1], out_36_8[1], out_36_9[1], 
               rep(out_10[1], length_101))

v_c_out_37 <- c(out_36_0[1], out_36_1[1], out_36_2[1], out_36_3[1], out_36_4[1], out_36_5[1], out_36_6[1], out_36_7[1], out_36_8[1], out_36_9[1], 
               rep(out_10[1], length_101))

v_c_out_term <- c(out_36_0[1], out_36_1[1], out_36_2[1], out_36_3[1], out_36_4[1], out_36_5[1], out_36_6[1], out_36_7[1], out_36_8[1], out_36_9[1], 
                 rep(out_10[1], length_101))

# Vector of total costs by GA
v_c_28 <- v_c_in_28 + v_c_out_28
v_c_32 <- v_c_in_32 + v_c_out_32
v_c_34 <- v_c_in_34 + v_c_out_34
v_c_37 <- v_c_in_37 + v_c_out_37
v_c_term <- v_c_in_term + v_c_out_term

## Plot for costs
# Dataframe for the cost vectors
df_costs_plot <- data.frame(age = c(0:100),
                            c_28 = v_c_28,
                            c_32 = v_c_32,
                            c_34 = v_c_34,
                            c_37 = v_c_37,
                            c_term = v_c_term)
# Plot
ggplot(data = df_costs_plot, aes(x = age)) +
  geom_line(aes(y = c_28), color = "darkgreen") +
  geom_line(aes(y = c_32), color = "darkblue") +
  geom_line(aes(y = c_34), color = "darkred") +
  labs(x = "Age", 
       y = "Total cost",
       title = "Age-dependent costs (Johnston 2014)") +
  theme_bw()
```

### QALYs

Age-dependent HRQoL (HUI3 scores) estimates by gestational age at birth. Health utility index mark 3 data is available from the 2015/2016 Canadian Community Health Survey (CCHS). Data are available from 12 to 80+ years in age range categories. The HUI3 values are adjusted by weights reported in the CCHS data. Initial HUI3 values (i.e. at age 0) are sourced from a multilevel model reported by Saigal et al. (2016). Based on these HUI3 values, a non-linear curve was fitted to approximately represent the trajectory of HUI3 values through ages 0 to 100. 

SOURCES:

* CCHS 2015-2016
* Saigal J Peds 2016 (PMID: 27592095)
* Vohr Pediatrics 2005 (PMID: 16143580)
* Johnson Arch Dis Child Fetal Neonatal Ed 2015 (PMID: 25834170)

```{r}
## Baseline HUI3 by age (SOURCE IS CCHS 2015/16 via Odesi)
# For instructions on incorporating weights see "Data/CCHS_2015_2016_User_Guide.pdf (s.10.2.3)"
# Relevant columns maintained: Age category (DHHGAGE), HUI3 outcome (HUIDVHSI), individual record weight (WTS_M)
# Remove missing HUI3 data (i.e. 99.999 and 99.996)
cchs_hui3 <- read.csv("Data/CCHS_2015-2016_modified.csv", header = TRUE) %>%
  group_by(DHHGAGE) %>%
  rename(age_cat = DHHGAGE, hui3 = HUIDVHSI) %>%
  # Generate the weighted HUI3 value for a given individual (numerator)
  mutate(hui3_w = hui3 * WTS_M)

## Define age categories from CCHS data
age_cat <- data.frame(age_cat = c(1:16),
                      age_range = c("12-14", "15-17", "18-19", "20-24", "25-29", "30-34","35-39","40-44", "45-49", "50-54", "55-59", "60-64", "65-69", "70-74", "75-79", "80+"),
                      age_n = c(3, 3, 2, 5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 21))

# Term (baseline) initial value 
term_init <- 0.871 # SOURCE --> Saigal 2016

term_init_28 <- term_init
term_init_32 <- term_init
term_init_34 <- term_init
term_init_37 <- term_init
term_init_term <- term_init 

## Calculate HUI3 summary data by age group and extend to age 0
hui3_sum <- cchs_hui3 %>%
  summarize(n = n(),
            mean_hui3_raw = mean(hui3),
            se_hui3_raw = sd(hui3)/sqrt(n()),
            wgt_sum = sum(WTS_M),
            hui3_w_sum = sum(hui3_w)) %>%
  mutate(mean_hui3 = hui3_w_sum/wgt_sum) %>% 
  # Organize data into annual categories
  merge(age_cat) %>%
  uncount(age_n) %>%
  arrange(age_cat) %>%
  mutate(age = c(12:100)) %>%
  # Add ages 0-11 (not available in the CCHS data), added from Saigal data
  add_row(mean_hui3 = rep(term_init, 12), age = 0:11, age_range = "0-11", .before = TRUE)

# Term (baseline) change (age 100-0)
# SOURCE --> CCHS, Saigal 2016
term_change <- -0.211 

term_change_28 <- term_change 
term_change_32 <- term_change 
term_change_34 <- term_change 
term_change_37 <- term_change 
term_change_term <- term_change

# Effect of neurosensory impairment AND pre-maturity on HUI3 relative to TERM controls
pt_nsi_i_dif <- -0.264

# Proportion of patients with NSI
p_nsi_28 <- 0.45
p_nsi_32 <- 0.28
p_nsi_34 <- 0.05
p_nsi_37 <- 0.07
p_nsi_term <- 0.02

## Difference in initial hui3 from pre-maturity alone (i.e. not NSI)
pt_i_dif_28 <- -0.092 # Known from Saigal 2016
pt_i_dif_32 <- 0 # Unknown, assumed to be zero
pt_i_dif_34 <- 0
pt_i_dif_37 <- 0

## Function to produce an annual HRQoL trend based on:
# 1. Impact of prematurity and NSI on y-intercept (hui3 score at age = 0) 
# 2. proportion of patients in the age category with neurosensory impairment

hui3_calc <- function(term_init. = term_init,
                      term_change. = term_change,
                      pt_i_dif. = 0,                  # set at default to term
                      pt_nsi_i_dif. = pt_nsi_i_dif,   
                      p_nsi_term. = p_nsi_term,
                      p_nsi_GA.){
  
  ## 1. Calculate the HUI3 curve for pt and pt + NSI
  # Variable terms (y scale shift i.e. initial and final values)
  
  # Without NSI (a = final value, b = initial value)
  a_pt <- term_init. + term_change. + pt_i_dif.
  a_pt <- ifelse(a_pt < 0, 0, a_pt)
  b_pt <- term_init. + pt_i_dif. 
  
  # With NSI (a = final value, b = initial value)
  a_pt_nsi <- term_init. + term_change. + pt_nsi_i_dif.
  a_pt_nsi <- ifelse(a_pt_nsi < 0, 0, a_pt_nsi)
  b_pt_nsi <- term_init. + pt_nsi_i_dif. 
  
  # Constant terms
  age <- seq(from = 0, to = 100, by = 1)
  c <- 72
  d <- 15
  
  # Curves for pt and pt_nsi
  pt <- a_pt + (b_pt - a_pt)/(1 + (age/c)^d)
  pt_nsi <- a_pt_nsi + (b_pt_nsi - a_pt_nsi)/(1 + (age/c)^d)
  
  ## 2. Combine the two curves (pt and pt_nsi) using p_nsi as a weight term
  # Difference in p_nsi from term (i.e. baseline neurosensory impairment probability)
  p_nsi <- abs(p_nsi_GA. - p_nsi_term.)
  
  # New curve as a weighted average
  final_curve <- (pt_nsi * p_nsi) + (pt * (1 - p_nsi))
  
  return(final_curve)
}

# Calculation of age-dependent HUI3 values for each GA category
v_q_28 <- hui3_calc(term_init. = term_init,
                    term_change. = term_change,
                    pt_i_dif. = pt_i_dif_28,
                    pt_nsi_i_dif. = pt_nsi_i_dif,
                    p_nsi_term. = p_nsi_term,
                    p_nsi_GA. = p_nsi_28)

v_q_32 <- hui3_calc(term_init. = term_init,
                    term_change. = term_change,
                    pt_i_dif. = pt_i_dif_32,
                    pt_nsi_i_dif. = pt_nsi_i_dif,
                    p_nsi_term. = p_nsi_term,
                    p_nsi_GA. = p_nsi_32)

v_q_34 <- hui3_calc(term_init. = term_init,
                    term_change. = term_change,
                    pt_i_dif. = pt_i_dif_34,
                    pt_nsi_i_dif. = pt_nsi_i_dif,
                    p_nsi_term. = p_nsi_term,
                    p_nsi_GA. = p_nsi_34)

v_q_37 <- hui3_calc(term_init. = term_init,
                    term_change. = term_change,
                    pt_i_dif. = pt_i_dif_37,
                    pt_nsi_i_dif. = pt_nsi_i_dif,
                    p_nsi_term. = p_nsi_term,
                    p_nsi_GA. = p_nsi_37)

v_q_term <- hui3_calc(term_init. = term_init,
                    term_change. = term_change,
                    pt_i_dif. = 0,
                    pt_nsi_i_dif. = 0,
                    p_nsi_term. = p_nsi_term,
                    # Set to 0 to avoid cancelling out and making p_nsi_term = 0 in calc.
                    p_nsi_GA. = 0) 

# All HUI3 curves in a single dataframe
hui3_curves <- data.frame(age = hui3_sum$age,
                          v_q_28,
                          v_q_32,
                          v_q_34,
                          v_q_37,
                          v_q_term) %>%
  pivot_longer(cols = c(2:6), names_to = "category", values_to = "hui3")

# Plot of annual HUI3 scores by GA category
hui3_plot <- ggplot(data = hui3_curves, aes(x = age, y = hui3)) +
  geom_line(aes(color = category), alpha = 0.7) +
  geom_line(data = hui3_sum, aes(y = mean_hui3, color = "CCHS"), color = "black", linetype = "dashed") + 
  ylim(c(0, 1)) +
  labs(x = "Age", 
       y = "HUI3") +
  scale_color_discrete(name = "GA categories (weeks)",
                       labels = (v_GA_cat)) +
  theme_bw()

hui3_plot

#ggsave(plot = hui3_plot, filename = "hui3_plot.pdf", path = "Results/", width = 10, height = 7)
```
